\section{Confluence}

\input{figures/02/par_reduction.tex}

The proof of confluence follows the PLFA book \cite{plfa22.08}.
This strategy involves the common technique of defining a parallel reduction variant of the one-step reduction described in Figure~\ref{fig:reduction}.
Parallel reduction allows reduction steps to occur in any subexpression, but reductions that generate new redexes cannot be reduced in a single step.
Figure~\ref{fig:par-reduction} presents the inductive definition of parallel reduction.
In fact, it is possible to compute the resulting syntax after all possible redexes are contracted by a single parallel reduction step.
This is the \textit{reduction completion} (written $\parp{t}$).
The definition of reduction completion is shown in Figure~\ref{fig:par-triangle}.
Reduction completion enables the derivation of a triangle property for which confluence for parallel reduction is a consequence.
Confluence for multistep reduction is an immediate consequence of confluence for parallel reduction and logical equivalence between both reduction variants.

\input{figures/02/triangle.tex}

\begin{lemma}
    For any $t$, $t \parred t$
    \label{lem:a:par_refl}
\end{lemma}
\begin{proof}
    Straightforward by induction on $t$.
\end{proof}

\begin{lemma}
    If $s \betared t$ then $s \parred t$
    \label{lem:a:beta_implies_par_step}
\end{lemma}
\begin{proof}
    By induction on $s \betared t$.
    The projection and promotion cases are similar to the substitution and beta case and thus omitted.
    The second structural binder reduction case is omitted.

    $\text{Case: }\begin{array}{c} \app{(\abs{\lambda_m}{x}{A}{b})}{m}{t} \betared [x := t]b \end{array}$
    \begin{proofcase}
        By Lemma~\ref{lem:a:par_refl}: $t \parred t$ and $b \parred b$.
        Applying the \textsc{ParBeta} rule concludes the case.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \app{\psi(\text{refl}(z; Z), u, v; A, P)}{\omega}{b} \betared b \end{array}$
    \begin{proofcase}
        Using Lemma~\ref{lem:a:par_refl}: $z \parred z$, $Z \parred Z$, $u \parred u$, $v \parred v$, $A \parred A$, $P \parred P$, and $b \parred b$.
        Applying the \textsc{ParSubst} rule concludes the case.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ConstructorReduction[*] \end{array}$
    \begin{proofcase}
        By the IH applied to $\D{1}$: $t_i \parred t_i^\prime$.
        Note that there is only one subderivation.
        For all $j \neq i$ $t_j \parred t_j$ by Lemma~\ref{lem:a:par_refl}.
        Using the \textsc{ParCtor} rule concludes the case.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \BinderOneReduction[*] \end{array}$
    \begin{proofcase}
        Applying the IH to $\D{1}$ yields $t_1 \parred t_1^\prime$.
        By Lemma~\ref{lem:a:par_refl}: $t_2 \parred t_2$.
        Using the \textsc{ParBind} rule concludes the case.
    \end{proofcase}
\end{proof}

\begin{lemma}
    If $s \betastar t$ then $s \parstar t$
    \label{lem:a:beta_implies_par}
\end{lemma}
\begin{proof}
    By induction on $s \betastar t$ applying Lemma~\ref{lem:a:beta_implies_par_step} in the inductive case.
\end{proof}

\begin{lemma}
    If $s \parred t$ then $s \betastar t$
    \label{lem:a:par_implies_beta_step}
\end{lemma}
\begin{proof}
    By induction on $s \parred t$.
    The projection, promotion, and substitution cases are similar to the beta case with the only difference being applying the associated rule.

    $\text{Case: }\begin{array}{c} \ParVarReduction[*] \end{array}$
    \begin{proofcase}
        By reflexivity of reduction.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParConstructorReduction[*] \end{array}$
    \begin{proofcase}
        By the IH applied to each $\D{i}$: $t_i \betastar t_i^\prime$ for all $i$.
        Applying Lemma~\ref{lem:2:beta_par} concludes the case.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParBinderReduction[*] \end{array}$
    \begin{proofcase}
        As the previous case, the IH yields $t_1 \betastar t_1$ and $t_2 \betastar t_2^\prime$.
        Again using Lemma~\ref{lem:2:beta_par} concludes the case.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParBetaReduction[*] \end{array}$
    \begin{proofcase}
        Applying the IH to all available derivations and using Lemma~\ref{lem:2:beta_par} gives $\app{(\abs{\lambda_m}{x}{t_1}{t_2})}{m}{t_3} \betastar \app{(\abs{\lambda_m}{x}{t_1^\prime}{t_2^\prime})}{m}{t_3^\prime}$.
        Applying the beta rule of reduction with transitivity concludes the case.
    \end{proofcase}
\end{proof}

\begin{lemma}
    If $s \parstar t$ then $s \betastar t$
    \label{lem:a:par_implies_beta}
\end{lemma}
\begin{proof}
    By induction on $s \parstar t$ applying Lemma~\ref{lem:a:par_implies_beta_step} in the inductive case.
\end{proof}

\begin{lemma}
    If $s \parred s^\prime$ and $t \parred t^\prime$ then $[x := s]t \parred [x := s^\prime]t^\prime$
    \label{lem:a:par_subst}
\end{lemma}
\begin{proof}
    By induction on $t \parred t^\prime$.
    The second projection case is omitted because it is the same as the first projection case.

    $\text{Case: }\begin{array}{c} \ParVarReduction[*] \end{array}$
    \begin{proofcase}
        Rename to $y$.
        If $x = y$ then $s \parred s^\prime$ which is a premise.
        If $x \neq y$ then no substitution is performed and $y_K \parred y_K$.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParConstructorReduction[*] \end{array}$
    \begin{proofcase}
        Applying the IH to $\D{i}$ yields $[x := s]t_i \parred [x := s^\prime]t_i^\prime$ for all $i$.
        Unfolding substitution for $\mathfrak{c}$ and applying the \textsc{ParCtor} rule concludes the case.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParBinderReduction[*] \end{array}$
    \begin{proofcase}
        As above the IH gives $[x := s]t_i \parred [x := s^\prime]t_i^\prime$ for $i = 1$ and $i = 2$.
        Unfolding substitution for $\mathfrak{b}$ and applying the \textsc{ParBind} rule concludes.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParBetaReduction[*] \end{array}$
    \begin{proofcase}
        By the IH: $[x := s]t_i \parred [x := s^\prime]t_i^\prime$ for $i = 1, 2, 3$.
        The \textsc{ParBeta} rule gives the following:
        $[x := s]\app{(\abs{\lambda_m}{y}{t_1}{t_2})}{m}{t_3} = \app{(\abs{\lambda_m}{y}{[x := s]t_1}{[x := s]t_2})}{m}{[x := s]t_3} \parred [y := t_3^\prime][x := s^\prime]t_2^\prime$.
        Note that $y$ is bound and thus not a free variable in $s^\prime$ and, moreover, by implicit renaming $x \neq y$.
        Thus, by Lemma~\ref{lem:2:subst_commute} $[y := t_3^\prime][x := s^\prime]t_2^\prime = [x := s^\prime][y := t_3^\prime]t_2^\prime$.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParSubstReduction[*] \end{array}$
    \begin{proofcase}
        By the IH: $[x := s]t_i \parred [x := s^\prime]t_i^\prime$ for $i = 1, 2$.
        The \textsc{ParSubst} rule gives:
        $[x := s](\app{\psi(\text{refl}(t_1; t_2), t_3, t_4; t_5, t_6)}{\omega}{t_7}) = \app{\psi(\text{refl}([x := s]t_1; [x := s]t_2), [x := s]t_3, [x := s]t_4; [x := s]t_5, [x := s]t_6)}{\omega}{[x := s]t_7} \parred [x := s^\prime]t_7^\prime$.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParFstReduction[*] \end{array}$
    \begin{proofcase}
        By the IH: $[x := s]t_i \parred [x := s^\prime]t_i^\prime$ for $i = 1, 2, 3$.
        The \textsc{ParFst} rule gives:
        $[x := s][t_1, t_2; t_3].1 = [[x := s]t_1, [x := s]t_2; [x := s]t_3].1 \parred [x := s^\prime]t_1^\prime$.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParPrmReduction[*] \end{array}$
    \begin{proofcase}
        By the IH: $[x := s]t_i \parred [x := s^\prime]t_i^\prime$ for $i = 1, 2, 3$.
        The \textsc{ParFst} rule gives:
        $[x := s]\vartheta(\text{refl}(t_1; t_2), t_3, t_4; t_5) = \vartheta(\text{refl}([x := s]t_1; [x := s]t_2), [x := s]t_3, [x := s]t_4; [x := s]t_5)  \parred \text{refl}([x := s^\prime]t_3^\prime; [x := s^\prime]t_5^\prime) = [x := s^\prime]\text{refl}(t_3^\prime; t_5^\prime)$.
    \end{proofcase}
\end{proof}

\noindent \begin{minipage}{.8\textwidth}
    The triangle property of parallel reduction is used to complete the set of possible contractible redexes.
    Thus, if syntax $s \parred t$ where $t$ is only partially reduced then both $s$ and $t$ may be completed to $\parp{s}$.
    To the right the situation is visually depicted.
    Note that the triangle property is ``half'' of the diamond property.
    Indeed, if $s \parred t^\prime$ then $t^\prime \parred \parp{s}$.
    As a consequence of the triangle property, parallel reduction trivially has the diamond property.
\end{minipage}\hfill
\begin{minipage}{.2\textwidth}
    \[\begin{tikzcd}
        s \\
        & t \\
        {\parp{s}}
        \arrow[Rightarrow, scaling nfold=3, from=1-1, to=3-1]
        \arrow[Rightarrow, scaling nfold=3, from=1-1, to=2-2]
        \arrow[Rightarrow, scaling nfold=3, from=2-2, to=3-1]
    \end{tikzcd}\]
\end{minipage}

\begin{lemma}[Parallel Triangle]
    If $s \parred t$ then $t \parred \parp{s}$
    \label{lem:a:par_triangle}
\end{lemma}
\begin{proof}
    By induction on $s \parred t$.
    The second projection case is omitted.

    $\text{Case: }\begin{array}{c} \ParVarReduction[*] \end{array}$
    \begin{proofcase}
        Have $\parp{x_K} = x_K$.
        Thus, this case is trivial.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParConstructorReduction[*] \end{array}$
    \begin{proofcase}
        By the IH applied to $\D{i}$: $t_i^\prime \parred \parp{t_i}$ for all $i$.
        Proceed by cases of $\parp{\mathfrak{c}(\kappa, t_1, \ldots t_{\mathfrak{a}(\kappa)})}$.
        The second projection case is omitted because it is the same as the first projection case.

        $\text{Case: }\begin{array}{c} \parp{\app{(\abs{\lambda_m}{x}{t_1}{t_2})}{m}{t_3}} = [x := \parp{t_3}]\parp{t_2} \end{array}$
        \begin{proofcase}
            Note that $\mathfrak{c}(\kappa, t_1^\prime, \ldots t^\prime_{\mathfrak{a}(\kappa)}) = \app{(\abs{\lambda_m}{x}{t_1^\prime}{t_2^\prime})}{m}{t_3^\prime}$.
            Using the \textsc{ParBeta} rule yields $\app{(\abs{\lambda_m}{x}{t_1^\prime}{t_2^\prime})}{m}{t_3^\prime} \parred [x := \parp{t_3}]\parp{t_2}$.
        \end{proofcase}

        $\text{Case: }\begin{array}{c} \parp{\psi(\app{\text{refl}(t_1; t_2), t_3, t_4; t_5, t_6)}{\omega}{t_7}} = \parp{t_7} \end{array}$
        \begin{proofcase}
            Note that $\mathfrak{c}(\kappa, t_1^\prime, \ldots t^\prime_{\mathfrak{a}(\kappa)}) = \psi(\app{\text{refl}(t_1^\prime; t_2^\prime), t_3^\prime, t_4^\prime; t_5^\prime, t_6^\prime)}{\omega}{t_7^\prime}$.
            Using the \textsc{ParSubst} rule yields $\psi(\app{\text{refl}(t_1^\prime; t_2^\prime), t_3^\prime, t_4^\prime; t_5^\prime, t_6^\prime)}{\omega}{t_7^\prime} \parred \parp{t_7}$.
        \end{proofcase}

        $\text{Case: }\begin{array}{c} \parp{[t_1, t_2; t_3].1} = \parp{t_1} \end{array}$
        \begin{proofcase}
            Note that $\mathfrak{c}(\kappa, t_1^\prime, \ldots t^\prime_{\mathfrak{a}(\kappa)}) = [t_1^\prime, t_2^\prime; t_3^\prime].1$.
            Using the \textsc{ParFst} rule yields $[t_1^\prime, t_2^\prime; t_3^\prime].1 \parred \parp{t_1}$.
        \end{proofcase}

        $\text{Case: }\begin{array}{c} \parp{\vartheta(\text{refl}(t_1; t_2), t_3, t_4; t_5)} = \prefl(\parp{t_3}; \parp{t_5}) \end{array}$
        \begin{proofcase}
            Note that $\mathfrak{c}(\kappa, t_1^\prime, \ldots t^\prime_{\mathfrak{a}(\kappa)}) = \vartheta(\text{refl}(t_1^\prime; t_2^\prime), t_3^\prime, t_4^\prime; t_5^\prime)$.
            Using the \textsc{ParPrmFst} rule yields $\vartheta(\text{refl}(t_1^\prime; t_2^\prime), t_3^\prime, t_4^\prime; t_5^\prime) \parred \prefl(\parp{t_3}; \parp{t_5})$.
        \end{proofcase}

        $\text{Case: }\begin{array}{c} \parp{\mathfrak{c}(\kappa, t_1, \ldots t_{\mathfrak{a}(\kappa)})} = \mathfrak{c}(\kappa, \parp{t_1}, \ldots \parp{t_{\mathfrak{a}(\kappa)}}) \end{array}$
        \begin{proofcase}
            Using the \textsc{ParCtor} rule concludes the case.
        \end{proofcase}
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParBinderReduction[*] \end{array}$
    \begin{proofcase}
        Note that $\parp{\mathfrak{b}(\kappa, (x : t_1), t_2)} = \mathfrak{b}(\kappa, (x : \parp{t_1}), \parp{t_2})$.
        By the IH applied to $\D{i}$: $t_i^\prime \parred \parp{t_i}$ for $i = 1, 2$.
        Thus, by the \textsc{ParBind} rule $\mathfrak{b}(\kappa, (x : t_1^\prime), t_2^\prime) \parred \mathfrak{b}(\kappa, (x : \parp{t_1}), \parp{t_2})$.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParBetaReduction[*] \end{array}$
    \begin{proofcase}
        Note that $\parp{\app{(\abs{\lambda_m}{x}{t_1}{t_2})}{m}{t_3}} = [x := \parp{t_3}]\parp{t_2}$.
        By the IH applied to $\D{i}$: $t_i^\prime \parred \parp{t_i}$ for $i = 1, 2, 3$.
        Thus, by Lemma~\ref{lem:a:par_subst} $[x := t_3^\prime]t_2^\prime \parred [x := \parp{t_3}]\parp{t_2}$.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParSubstReduction[*] \end{array}$
    \begin{proofcase}
        Note that $\parp{\app{\psi(\text{refl}(t_1; t_2), t_3, t_4; t_5, t_6)}{\omega}{t_7}} = \parp{t_7}$.
        By the IH applied to $\D{i}$: $t_i^\prime \parred \parp{t_i}$ for $i = 1$ through $i = 7$.
        Applying the \textsc{ParBind} rule yields $t_7^\prime \parred \parp{t_7}$.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParFstReduction[*] \end{array}$
    \begin{proofcase}
        Note that $\parp{[t_1, t_2; t_3].1} = \parp{t_1}$.
        By the IH applied to $\D{i}$: $t_i^\prime \parred \parp{t_i}$ for $i = 1, 2, 3$.
        Thus, $t_1^\prime \parred \parp{t_1}$.
    \end{proofcase}

    $\text{Case: }\begin{array}{c} \ParPrmReduction[*] \end{array}$
    \begin{proofcase}
        Note that $\parp{\vartheta_1(\text{refl}(t_1; t_2), t_3, t_4; t_5)} \parred \prefl(\parp{t_3}; \parp{t_5})$.
        By the IH applied to $\D{i}$: $t_i^\prime \parred \parp{t_i}$ for $i = 1$ through $i = 5$.
        Thus, $\prefl(t_3^\prime;t_5^\prime) \parred \prefl(\parp{t_3};\parp{t_5})$ by the \textsc{ParCtor} rule and Lemma~\ref{lem:a:par_refl}.
    \end{proofcase}
\end{proof}

\begin{lemma}[Parallel Strip]
    \label{lem:a:par_strip}
    If $s \parred t_1$ and $s \parstar t_2$ then $\exists\ t$ such that $t_1 \parstar t$ and $t_2 \parred t$
\end{lemma}
\begin{proof}
    By induction on $s \parstar t_2$, pick $t = t_1$ for the reflexivity case.
    Consider the transitivity case, $\exists\ z_1$ such that $s \parred z_1$ and $z_1 \parstar t_2$.
    Applying Lemma~\ref{lem:a:par_triangle} to $s \parred z_1$ yields $z_1 \parred \parp{s}$.
    By the IH with $z_1 \parred \parp{s}$: $\exists\ z_2$ such that $\parp{s} \parstar z_2$ and $t_2 \parred z_2$.
    Using Lemma~\ref{lem:a:par_triangle} again on $s \parred t_1$ yields $t_1 \parred \parp{s}$.
    Now by transitivity $t_1 \parstar z_2$.
\end{proof}

\begin{lemma}[Parallel Confluence]
    \label{lem:a:par_confluence}
    If $s \parstar t_1$ and $s \parstar t_2$ then $\exists\ t$ such that $t_1 \parstar t$ and $t_2 \parstar t$
\end{lemma}
\begin{proof}
    By induction on $s \parstar t_1$, pick $t = t_2$ for the reflexivity case.
    Consider the transitivity case, $\exists\ z_1$ such that $s \parred z_1$ and $z_1 \parstar t_1$.
    By Lemma~\ref{lem:a:par_strip} applied with $s \parred z_1$ and $s \parstar t_2$ yields $\exists\ z_2$ such that $z_1 \parstar z_2$ and $t_2 \parred z_2$.
    Using the IH with $z_1 \parred z_2$ gives $\exists\ z_3$ such that $t_1 \parstar z_3$ and $z_2 \parstar z_3$.
    By transitivity $t_2 \parstar z_3$.
\end{proof}

\begin{lemma}[Confluence]
    \label{lem:2:confluence}
    If $s \betastar t_1$ and $s \betastar t_2$ then $\exists\ t$ such that $t_1 \betastar t$ and $t_2 \betastar t$
\end{lemma}
\begin{proof}
    By Lemma~\ref{lem:a:beta_implies_par} applied twice: $s \parstar t_1$ and $s \parstar t_2$.
    Now by parallel confluence (Lemma~\ref{lem:a:par_confluence}) $\exists\ t$ such that $t_1 \parstar t$ and $t_2 \parstar t$.
    Finally, two applications of Lemma~\ref{lem:a:par_implies_beta} conclude the proof.
\end{proof}

As with F$^\omega$ the important consequence of confluence is that convertibility of reduction is an equivalence relation.
However, this is \textit{not} the conversion relation that will be used in the inference judgment.
Thus, while important, it is still only a stepping stone to showing judgmental conversion is transitive.

\begin{theorem}
    \label{lem:2:beta_conv_equivalence}
    For any $s$ and $t$ the relation $s \betaconv t$ is an equivalence.
\end{theorem}
\begin{proof}
    Reflexivity is immediate because $s \betastar s$.
    Symmetry is also immediate because if $s \betaconv t$ then $\exists\ z$ such that $s \betastar z$ and $t \betastar z$, but logical conjunction is commutative.
    Transitivity is a consequence of confluence, see Theorem~\ref{thm:1:trans}.
\end{proof}

Additionally, there is a final useful fact about reduction conversion that is occasionally used throughout the rest of this work.
That is, like reduction, conversion of subexpressions yields conversion of the entire term.

\begin{lemma}
    \label{lem:2:conv_congr}
    If $t_i \betaconv t_i^\prime$ for any $i$ then,
    \begin{enumerate}
        \item $\mathfrak{b}(\kappa, (x : t_1), t_2) \betaconv \mathfrak{b}(\kappa, (x : t_1^\prime), t_2^\prime)$
        \item $\mathfrak{c}(\kappa, t_1,\ldots,t_{\mathfrak{a}(\kappa)}) \betaconv \mathfrak{c}(\kappa, t_1^\prime,\ldots,t_{\mathfrak{a}(\kappa)}^\prime)$
    \end{enumerate}
\end{lemma}
\begin{proof}
    By Lemma~\ref{lem:2:beta_par} applied on both sides.
\end{proof}
